A Streaming Algorithm for 2-Center with Outliers in High Dimensions

### 2015-07-12 outlier.tex
Camera-Ready

\newtheorem{problem}{Problem}
\newtheorem{definition}{Definition}

@@

\title{A Streaming Algorithm for 2-Center with Outliers in High Dimensions}
%\title{Computing 2-Center with Outliers on High-Dimensional Data Streams}
\author{Behnam~Hatami\thanks{Department of Computer Engineering, 
	Sharif University of Technology.
	{\tt bhatami@ce.sharif.edu}}
	\and 
	Hamid~Zarrabi-Zadeh\thanks{Department of Computer Engineering, 
	Sharif University of Technology.
	{\tt zarrabi@sharif.edu}}
}

@@

To obtain our result, we have used a combination of several ideas
including parallelization construction, far/close ball separation, centerpoint theorem,
and keeping a lower bound on the radius and distance of the optimal balls.
We have also utilized ideas used in~\cite{kim2014improved} 
for the 2-center problem with no outliers.
%However, our algorithm is much detailed and uses many new ingredients.
However, our problem is much harder here, % handling outliers makes the problem much harder,
as we not only need to find balls of minimum radius, 
but we also need to decide which subset of points to cluster.
This is in particular more challenging in the streaming model,
where we only have a single pass over the input, and we must decide on the fly
which point is an outlier, 
and which one can be safely ignored as a non-outlier point,
%using only a limited amount of storage.
to comply with the working space restriction enforced by the model.

@@

\begin{proof}
Suppose by way of contradiction that $S \cap B_2^*$ is empty. 
Since $\card{S} = z + 1$, there is at least one point in $S$
which is not outlier, and hence, it is in $B_1^*$.
Let $q$ be the point in $S \cap B_1^*$ furthest from~$p$. 
Consider $B(p, \len{pq})$. It is clear that $B$ covers $P \setminus S$. 
Since $p, q \in B_1^*$, then $\len{pq}$ is at most $2r^*$. 
Thus, by Observation~\ref{obs:intersection}, $B_2^* \cap B = \emptyset$. 
Therefore, $B_2^* \cap P = \emptyset$, and hence,
 $B_2^*$ is empty, which contradicts the optimality of the solution.
\end{proof}

@@

\begin{algorithm}[t]
\caption{\sc 2-CenterCase1$(P)$} 
\algsetup{indent=1.5em}
\label{alg:case-1}
\begin{algorithmic}[1]
\STATE $c_1 \gets p_1$, $\rc  \gets 0$, $\delta \gets 0$
\FORALL{$p \in P$}
	\IF {\NOT (\Call{AddToB$_1$}{$p$} \OR \Call{AddToB$_2$}{$p$})}
		\STATE add $p$ to Buffer
		\WHILE{$\card{\mbox{Buffer}} > z$} \label{step:buffer-overflow}
			\IF{$\card{B_2} \ge \dz$}
				\STATE $B_1 \gets B_1 \cup B_2$, $B_2 \gets \emptyset$
			\ELSIF{$c_2$ is set}
				\STATE $B_1 \gets B_1 \cup \set{c_2}$ %, $B_u \gets B_u \cup \set{c_2}$
				%\STATE remove $c_2$ from $B_2$
			\ENDIF
			\STATE $T \gets \mbox{Buffer} \cup B_2 \setminus \set{c_2}$
			\STATE $B_2 \gets \emptyset$ 
%			\IF{first iteration of while}
%				\FOR{$p \in T$}
%					\IF{\Call{AddToB$_1$}{$p$}}
%						\STATE remove $p$ from $T$
%					\ENDIF
%				\ENDFOR
%			\ENDIF
			\STATE $c_2 \gets$ (z+1)-furthest point from $c_1$ in $T$ 
			\STATE $\rc  \gets \radius{c_2}$
			\FOR{$p \in T$}
				\STATE \Call{AddToB$_2$}{$p$}
			\ENDFOR				
			\STATE $\mbox{Buffer} \gets T \setminus B_2$
		\ENDWHILE
	\ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

Invariant (c): 
We first claim that if $c_2$ is set, then
$B(c_p, \radius{c_p}) \subseteq B_2(c_2, r)$.
If $|B_2| < \dz$, then $c_p = c_2$ and $r = \radius{c_2}$, 
and hence, $B_2 = B(c_p, \radius{c_p})$. 
When the size of $B_2$ reaches $\dz$, 
the central point $\cp$ moves to the centerpoint of $B_2$,
and $\rc$ is increased by a factor of $(2 + \frac{2}{\alpha})$. 
Because the centerpoint of $B_2$ locates in $B_2$, then $c_p \in B(c_2, \radius{c_2})$.
Thus, if $|B_2| \ge \dz$ then $\len{c_2 c_p} \leq \radius{c_2}$, and hence
$$
	\radius{c_p} \leq \frac{2 (\len{c_1 c_2} + \len{c_2 c_p})}{\alpha} \leq \radius{c_2}(1 + \frac{2}{\alpha}).
$$ 
Therefore, $B(c_p, \radius{c_p}) \subseteq B_2(c_2, \radius{c_2}(2 + \frac{2}{\alpha}))$.

@@

The second case is in \Cref{alg:case-1}, when the buffer overflows and $B_2$ is non-empty.
The algorithm takes one of the following actions depending on the size of $B_2$.
If $|B_2| < \dz $, then $c_2 = c_p$. \Cref{alg:case-1} add $c_2$ to $B_1$.
By contradiction assume that $c_p \in B_2^*$, 
then by invariants (b) and (d1), 
$2r^* \leq r \leq \delta/2 < \delta^*/2$. 
Therefore, by Lemma \ref{lem:2r}, there must be at most $z$ points outside $B_1 \cup B_2$,
which contradicts the overflow of the buffer. 
If $|B_2| \geq \dz$, then $c_p$ is the centerpoint of the first $\dz$ points currently in $B_2$. 
In this case, we add all points of $B_2$ to $B_1$. 
By invariant (b), $r \leq \delta/2 < \delta^*/2$. 
Therefore, By Lemma \ref{lem:center-point}, $c_p \in B_1^*$ or $c_p \in B_2^*$. 
Suppose by way of contradiction that $c_p \in B_2^*$. 
In this case, by invariant (d1) and the claim we proved first,
$B_2$ covers $B(c_p, \radius{c_p})$ and $2r^* \leq r$. 
Therefore, Similar to the previous part, it contradicts the overflow of the buffer.

@@

Invariant (d3): By invariants (c) and (d1), $2r^* \leq r \leq \delta/2 < \delta^* /2$. 
Therefore, by Lemma \ref{lem:2r}, all points outside $B_1 \cup B_2$ are outlier.

\end{proof}

@@

%Our algorithm for solving the problem in this case is based on two main ingredients.
Our idea in this section is to carefully adopt
the algorithm of Kim and Ahn~\cite{kim2014improved},
originally designed for maintaining an approximate 2-center.
To avoid duplication, we just sketch the main steps of their algorithm,
%without entering the details, 
and explain our modifications to it. 
Kim and Ahn's algorithm, which we refer to as the KA algorithm,
has 9 different states, shown in~\Cref{fig:dag}.
Depending on the points arrived so far, the algorithm is in one of the states.
%In each step one or many of the states can be valid and their algorithm considers all of them in parallel. 
In each state, the algorithm keeps at most two balls, as a candidate solution. 
A transition between the states occurs whenever a point not covered by any of the two balls arrive.

@@

\begin{figure}
\centering
\begin{tikzpicture}
  [scale=.75,auto=left,every node/.style={circle,fill=blue!15}]
 \tikzset{edge/.style = {->,> = latex'}}
  \node (1) at (1,2.5) {1};
  \node (2a) at (3,1)  {2a};
  \node (3a) at (3,4)  {3a};
  \node (2b) at (5,0)  {2b};
  \node (3b) at (5,2)  {3b};
  \node (3d) at (5,4)  {3d};
  \node (2c) at (7,0)  {2c};
  \node (3c) at (7,2)  {3c};
  \node (3e) at (7,4)  {3e};
  %\node (E) at (9, 2) {E};

@@

\begin{algorithm}[t]
\caption{\sc 2-CenterCase2$(P, z, r)$} 
\algsetup{indent=1.5em}
\label{alg:smallc}
\begin{algorithmic}[1]
	\STATE solutions $\gets \set{}$
	\FORALL {$(n_1, n_2, n_3, n_4)$ such that $\sum n_i=z$}
		%\FORALL {path $\pi$ in the transition graph}
		\FORALL {$\pi \in \set{1, 2, 3}$}
			\STATE $\counter_i \gets 0$, for $i = 1, \ldots, 4$ % $1 \le i \le 4$
			\STATE $B_1 \gets B(p_1, r)$, $B_2 \gets \emptyset$
			\STATE $\level \gets 1$ \COMMENT{$\level$ represents current level}
			\FORALL {$p \in P$}
				\IF {$p \not\in B_1 \cup B_2$}
					\STATE $\counter_\level \gets \counter_\level + 1$
					\IF {$\counter_\level > n_\level$}
						\STATE $\level \gets \level+1$
						\STATE $(B_1, B_2) \gets \CA.\Insert(p, \pi)$
					\ENDIF
				\ENDIF
			\ENDFOR
			\IF {$\level \le 4$}
				\STATE add $\max \set{r(B_1), r(B_2)}$ to solutions
				%\STATE add $(B_1, B_2)$ to solutions
			\ENDIF
		\ENDFOR
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
\end{algorithmic}
\end{algorithm}

@@

\begin{theorem} \label{thm:close}
	For $1.2r^* \le r < (1.2 + {2 \over 3}\eps)r^*$ and $\delta^* \le \alpha r^*$,
	\Cref{alg:smallc} computes a $(1.8+\eps)$-approximation to 
	the 2-center with $z$ outliers 
	in $\O (dz^3)$ space and $\O (dz^3)$ update time.
\end{theorem}

@@

In this paper, we presented a $(1.8 + \eps)$-approximation streaming algorithm for 2-center problem with outliers in Euclidean space.
It improves the previous $(4+\eps)$-approximation algorithm available for the problem 
due to Charikar~\etal~\cite{mccutchen2008streaming}. 
Finding better approximation factor or space complexity
is an interesting problem that remains open.
It is also interesting to see if the ideas in this paper 
can be extended to the $k$-center problem with outliers in the data stream model,
%for any $k$, or 
even for small values of $k \ge 3$.

@@

%---------------------------- Appendix ------------------------

@@

\begin{proof}
	By invariants (a) and (b), we know that $r \leq \delta/2 < \delta^*/2$. 
	By Lemma~\ref{lem:center-point}, $c_p \in B_1^* \cap B_2^*$. 
	Since $c_p \not \in B_2^*$, then $c_p \in B_1^*$. 
	On the other hand, by Observation~\ref{obs:intersection}, 
	$B_2$ intersect at most one of $B_1^*$ and $B_2^*$. 
	Therefore, $B_2 \cap B_2^* = \emptyset$.
\end{proof}

\noindent
Our query algorithm works as follows. 
For each candidate point $c \in C$, \Cref{alg:query} 
constructs $B'_1(c_1, \max \{\delta, \radius{c} \})$ and $B'_2(c, \radius{c})$. 
If the candidate $c$ equals the current $c_2$, then, $B_1 = B'_1$. 
Since $\radius{c_2} \leq r \leq \delta/2$ by invariant (b) and $B'_2 \subset B_2$,
we do not need to construct any new set. 
For $c \neq c_2$, we know that $B_1 \subset B'_1$, and hence,
we just need to see which points in $\Buffer \cap B_2$ are located in $B'_1$.
When $|B_2| \geq \dz$ and $c \neq c_2$, then it means that assumed $c_p \not \in B_2^*$. 
Therefore, by Lemma \ref{lem:B2dz}, $B_2$ can be added to $B'_1$ without contradicting invariant (c). 
So in this case, we just need to see which points of $\Buffer$ must be added to $B'_1$. 
\Cref{alg:query} use functions \textproc{AddToB$'_1$} and \textproc{AddToB$'_2$} for adding a point to $B'_1$ and $B'_2$ respectively. 
The definition of these functions are the same as \textproc{AddToB$_1$} and \textproc{AddToB$_2$},
except that, they add point to $B'_i$ instead of $B_i$ ($i = 1, 2$).

Since \Cref{alg:query} considers all valid candidates for $c$, 
at least for one $c^* \in C$, we have $c^* \in B_2^*$. 
We call the corresponding $B'_1$ and $B'_2$ by $B''_1$ and $B''_2$. 
Since by Observation \ref{obs:c+4},
$1 \leq \frac{\len{c_1 c^*}}{\alpha^*} < \frac{\len {c_1 c^*}}{\alpha r^*}$, 
we have $2r^* \leq \radius{c^*}$, and therefore,
$B^*_1 \subset B''_1$ and $B^*_2 \subset B''_2$ by Lemma~\ref{lem:2r}.
On the other hand, since the distance of the new points added to $B''_1$ compare to $B_1$ is less than $\len{c_1 p} \leq \max \{\delta, \radius{c^*} \}$, 
we have by invariant~(c) that  $B''_1 \cap B_2^* = \emptyset$.
As a result, $B''_1$ (resp., $B''_2$) completely covers $B_1^*$  (resp., $B_2^*$), 
and points in $\Buffer$ are certainly outlier.
The only unknown part is that \Cref{alg:query} does not know how many outliers are in $B''_1$ and $B''_2$.
Therefore, \Cref{alg:compute-min} tries all possible cases and choose the one with the minimum radius.

@@

\REM{ }% to be added in the camera-ready version
\paragraph{Acknowledgement}
The authors would like to thank Kiana Ehsani and Sahand Mozaffari 
for their thoughtful discussions, 
and for their help to improve the quality of this paper.

>>>>

\newtheorem{problem}{Problem}
\newtheorem{definition}{Definition}
\newtheorem{claim}{Claim}

@@

\title{A Streaming Algorithm for 2-Center with Outliers in High Dimensions}

\author{Behnam~Hatami\thanks{Department of Computer Engineering, 
	Sharif University of Technology, Tehran, Iran.
	{\tt bhatami@ce.sharif.edu}}
	\and 
	Hamid~Zarrabi-Zadeh\thanks{Department of Computer Engineering, 
	Sharif University of Technology, Tehran, Iran.
	{\tt zarrabi@sharif.edu}}
}

@@

To obtain our result, we have used a combination of several ideas
including parallelization, far/close ball separation, centerpoint theorem,
and keeping a lower/upper bound on the radius and distance of the optimal balls.
We have also utilized ideas used in~\cite{kim2014improved} 
for the 2-center problem with no outliers.
%However, our algorithm is much detailed and uses many new ingredients.
However, our problem is much harder here, % handling outliers makes the problem much harder,
as we not only need to find balls of minimum radius, 
but we also need to decide which subset of points to cluster.
This is in particular more challenging in the streaming model,
where we only have a single pass over the input, and we must decide on the fly
which point is an outlier, 
and which one can be safely ignored as a non-outlier point,
%using only a limited amount of storage.
to comply with the working space restriction enforced by the model.

@@

\begin{proof}
Suppose by way of contradiction that $S \cap B_2^*$ is empty. 
Since $\card{S} = z + 1$, there is at least one point in $S$
which is not outlier, and hence, it is in $B_1^*$.
Let $q$ be a point in $S \cap B_1^*$ furthest from~$p$. 
Consider the ball $B(p, \len{pq})$. 
%It is clear that $B$ covers $P \setminus S$. 
For any point $s \in P \setminus S$, 
we have $\len{ps} \leq \len{pq}$,
because $s \not\in S$ and $q \in S$. 
Therefore, $B$ covers $P \setminus S$.
Since $p, q \in B_1^*$, $\len{pq}$ is at most $2r^*$. 
Thus, by Observation~\ref{obs:intersection}, $B_2^* \cap B = \emptyset$. 
Therefore, $B_2^* \cap P = \emptyset$, and hence,
 $B_2^*$ is empty, which contradicts the optimality of the solution.
\end{proof}

@@

\begin{algorithm}[t]
\caption{\sc 2-Center-Separated$(P)$} 
\algsetup{indent=1.5em}
\label{alg:case-1}
\begin{algorithmic}[1]
\STATE $c_1 \gets p_1$, $\rc  \gets 0$, $\delta \gets 0$
\FORALL{$p \in P$}
	\IF {\NOT (\Call{AddToB$_1$}{$p$} \OR \Call{AddToB$_2$}{$p$})}
		\STATE add $p$ to Buffer
		\WHILE{$\card{\mbox{Buffer}} > z$} \label{step:buffer-overflow}
			\IF{$\card{B_2} \ge \dz$}
				\STATE $B_1 \gets B_1 \cup B_2$, $B_2 \gets \emptyset$
			\ELSIF{$c_2$ is set}
				\STATE $B_1 \gets B_1 \cup \set{c_2}$ %, $B_u \gets B_u \cup \set{c_2}$
				%\STATE remove $c_2$ from $B_2$
			\ENDIF
			\STATE $T \gets \mbox{Buffer} \cup B_2 \setminus \set{c_2}$
			\STATE $B_2 \gets \emptyset$ 
%			\IF{first iteration of while}
%				\FOR{$p \in T$}
%					\IF{\Call{AddToB$_1$}{$p$}}
%						\STATE remove $p$ from $T$
%					\ENDIF
%				\ENDFOR
%			\ENDIF
			\STATE $c_2 \gets$ (z+1)-furthest point from $c_1$ in $T$ 
			\STATE $\rc  \gets \radius{c_2}$
			\FOR{$p \in T$}
				\STATE \Call{AddToB$_2$}{$p$}
			\ENDFOR				
			\STATE $\mbox{Buffer} \gets T \setminus B_2$
		\ENDWHILE
	\ENDIF
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

Invariant (c): The proof is provided in \Cref{app:inv-c}. 
% due to space constraints.

\REM{
Invariant (c): 
We first claim that if $c_2$ is set, then
$B(c_p, \radius{c_p}) \subseteq B_2(c_2, r)$.
If $|B_2| < \dz$, then $c_p = c_2$ and $r = \radius{c_2}$, 
and hence, $B_2 = B(c_p, \radius{c_p})$. 
When the size of $B_2$ reaches $\dz$, 
the central point $\cp$ moves to the centerpoint of $B_2$,
and $\rc$ is increased by a factor of $(2 + \frac{2}{\alpha})$. 
Because the centerpoint of $B_2$ locates in $B_2$, then $c_p \in B(c_2, \radius{c_2})$.
Thus, if $|B_2| \ge \dz$ then $\len{c_2 c_p} \leq \radius{c_2}$, and hence
$$
	\radius{c_p} \leq \frac{2 (\len{c_1 c_2} + \len{c_2 c_p})}{\alpha} \leq \radius{c_2}(1 + \frac{2}{\alpha}).
$$ 
Therefore, $B(c_p, \radius{c_p}) \subseteq B_2(c_2, \radius{c_2}(2 + \frac{2}{\alpha}))$.

@@

The second case is in \Cref{alg:case-1}, when the buffer overflows and $B_2$ is non-empty.
The algorithm takes one of the following actions depending on the size of $B_2$.
If $|B_2| < \dz $, then $c_2 = c_p$. \Cref{alg:case-1} add $c_2$ to $B_1$.
By contradiction assume that $c_p \in B_2^*$, 
then by invariants (b) and (d1), 
$2r^* \leq r \leq \delta/2 < \delta^*/2$. 
Therefore, by Lemma \ref{lem:2r}, there must be at most $z$ points outside $B_1 \cup B_2$,
which contradicts the overflow of the buffer. 
If $|B_2| \geq \dz$, then $c_p$ is the centerpoint of the first $\dz$ points currently in $B_2$. 
In this case, we add all points of $B_2$ to $B_1$. 
By invariant (b), $r \leq \delta/2 < \delta^*/2$. 
Therefore, By Lemma \ref{lem:center-point}, $c_p \in B_1^*$ or $c_p \in B_2^*$. 
Suppose by way of contradiction that $c_p \in B_2^*$. 
In this case, by invariant (d1) and the claim we proved first,
$B_2$ covers $B(c_p, \radius{c_p})$ and $2r^* \leq r$. 
Therefore, Similar to the previous part, it contradicts the overflow of the buffer. 
}

@@

Invariant (d3): By invariants (c) and (d1), $2r^* \leq r \leq \delta/2 < \delta^* /2$. 
Therefore, by Lemma \ref{lem:2r}, all points outside $B_1 \cup B_2$ are outlier.
\end{proof}

@@

%Our algorithm for solving the problem in this case is based on two main ingredients.
Our idea in this section is to carefully adopt
the algorithm of Kim and Ahn~\cite{kim2014improved},
originally designed for maintaining an approximate 2-center.
To avoid duplication, we just sketch the main steps of their algorithm,
%without entering the details, 
and explain our modifications to it. 
Kim and Ahn's algorithm, which we refer to as the KA algorithm,
has 9 different states, shown in~\Cref{fig:dag}.
Depending on the points arrived so far, the algorithm is in one of the states.
%In each step one or many of the states can be valid and their algorithm considers all of them in parallel. 
In each state, the algorithm keeps at most two balls as a candidate solution. 
A transition between the states occurs whenever a point not covered by any of the two balls arrive.

@@

\begin{figure}
\centering
\begin{tikzpicture}
  [scale=.75,auto=left,every node/.style={circle,fill=blue!15}]
 \tikzset{edge/.style = {->,> = latex'}}
  \node (1) at (1,2.5) {1};
  \node (2a) at (3,4)  {2a};
  \node (2b) at (5,4)  {2b};
  \node (2c) at (7,4)  {2c};
  \node (3a) at (3,1)  {3a};
  \node (3b) at (5,2)  {3b};
  \node (3c) at (7,2)  {3c};
  \node (3d) at (5,0)  {3d};
  \node (3e) at (7,0)  {3e};
  %\node (E) at (9, 2) {E};

@@

\begin{algorithm}[t]
\caption{\sc 2-Center-Close$(P, z, r)$} 
\algsetup{indent=1.5em}
\label{alg:smallc}
\begin{algorithmic}[1]
	\STATE solutions $\gets \set{}$
	\FORALL {$(n_1, n_2, n_3, n_4)$ such that $\sum n_i=z$}
		%\FORALL {path $\pi$ in the transition graph}
		\FORALL {$\pi \in \set{1, 2, 3}$}
			\STATE $\counter_i \gets 0$, for $i = 1, \ldots, 4$ % $1 \le i \le 4$
			\STATE $B_1 \gets B(p_1, r)$, $B_2 \gets \emptyset$
			\STATE $\level \gets 1$ \COMMENT{$\level$ represents current level}
			\FORALL {$p \in P$}
				\IF {$p \not\in B_1 \cup B_2$}
					\STATE $\counter_\level \gets \counter_\level + 1$
					\IF {$\counter_\level > n_\level$}
						\STATE $\level \gets \level+1$
						\STATE $(B_1, B_2) \gets \CA.\Insert(p, \pi)$
					\ENDIF
				\ENDIF
			\ENDFOR
			\IF {$\level \le 4$}
				\STATE add $\max \set{r(B_1), r(B_2)}$ to solutions
				%\STATE add $(B_1, B_2)$ to solutions
			\ENDIF
		\ENDFOR
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
\end{algorithmic}
\end{algorithm}

@@

\begin{theorem} \label{thm:close}
	For $1.2r^* \le r < (1.2 + {2 \over 3}\eps)r^*$ and $\delta^* \le \alpha r^*$,
	\Cref{alg:smallc} computes a $(1.8+\eps)$-approximation to 
	the 2-center with $z$ outliers 
	in $\O (dz^3)$ space and $\O (dz^3)$ update time,
	assuming that the first point of the stream is not outlier. 
\end{theorem}

@@

In this paper, we presented a $(1.8 + \eps)$-approximation streaming algorithm for 2-center problem with outliers in Euclidean space.
It improves the previous $(4+\eps)$-approximation algorithm available for the problem 
due to McCutchen and Khuller~\cite{mccutchen2008streaming}. 
Finding better approximation factor or space complexity
is an interesting problem that remains open.
It is also interesting to see if the ideas in this paper 
can be extended to the $k$-center problem with outliers in the data stream model,
%for any $k$, or 
even for small values of $k \ge 3$.

\paragraph{Acknowledgement}
The authors would like to thank Kiana Ehsani and Sahand Mozaffari 
for their thoughtful discussions, and for their very helpful comments.
%and for their help to improve the quality of this paper.

@@

%---------------------------- Appendix ------------------------
\newpage

@@

\begin{proof}
	By invariants (a) and (b), we know that $r \leq \delta/2 < \delta^*/2$. 
	By Lemma~\ref{lem:center-point}, $c_p \in B_1^* \cap B_2^*$. 
	Since $c_p \not \in B_2^*$, we have $c_p \in B_1^*$. 
	On the other hand, by Observation~\ref{obs:intersection}, 
	$B_2$ intersect at most one of $B_1^*$ and $B_2^*$. 
	Therefore, $B_2 \cap B_2^* = \emptyset$.
\end{proof}

\noindent
Our query algorithm works as follows. 
For each candidate point $c \in C$, \Cref{alg:query} 
constructs $B'_1(c_1, \max \{\delta, \radius{c} \})$ and $B'_2(c, \radius{c})$. 
If the candidate $c$ equals the current $c_2$, then we have $B_1 = B'_1$. 
Since $\radius{c_2} \leq r \leq \delta/2$ by invariant (b), and $B'_2 \subset B_2$,
we do not need to construct any new set. 
For $c \neq c_2$, we know that $B_1 \subset B'_1$, and hence,
we only need to see which points in $\Buffer \cap B_2$ are inside $B'_1$.
When $|B_2| \geq \dz$ and $c \neq c_2$, then it means that $c_p \not \in B_2^*$. 
Therefore, by Lemma \ref{lem:B2dz}, $B_2$ can be added to $B'_1$ without violating invariant (c). 
So in this case, we just need to see which points of $\Buffer$ must be added to $B'_1$. 
\Cref{alg:query} uses functions \textproc{AddToB$'_1$} and \textproc{AddToB$'_2$} 
for adding a point to $B'_1$ and $B'_2$ respectively. 
These functions are the same as \textproc{AddToB$_1$} and \textproc{AddToB$_2$},
with the only exception that they add points to $B'_i$ instead of $B_i$, for $i = 1, 2$.

Since \Cref{alg:query} considers all valid candidates for $c$, 
at least for one $c^* \in C$, we have $c^* \in B_2^*$. 
We denote the corresponding $B'_1$ and $B'_2$ by $B''_1$ and $B''_2$. 
Since by Observation \ref{obs:c+4},
$1 \leq \frac{\len{c_1 c^*}}{\alpha^*} < \frac{\len {c_1 c^*}}{\alpha r^*}$, 
we have $2r^* \leq \radius{c^*}$, and hence
by Lemma~\ref{lem:2r}, $B^*_1 \subset B''_1$ and $B^*_2 \subset B''_2$.
On the other hand, since the distance of the new points added to $B''_1$ 
%compare to $B_1$ 
is less than $\len{c_1 p} \leq \max \{\delta, \radius{c^*} \}$, 
we have by invariant~(c) that  $B''_1 \cap B_2^* = \emptyset$.
As a result, $B''_1$ (resp., $B''_2$) completely covers $B_1^*$  (resp., $B_2^*$), 
and the points in $\Buffer$ are all outliers.
The only unknown part is that \Cref{alg:query} does not know how many outliers are in $B''_1$ and $B''_2$.
Therefore, \Cref{alg:compute-min} tries all possible cases and choose the one with the minimum radius.

@@

% --------------------------- Invariant (c) ----------------------

\section{Proof of Invariant (c)}
\label{app:inv-c}

Here, we provide a proof for Invariant (c).
%We first prove the following technical claim.
The following technical claim will be used in our proof.

\begin{claim}
	\label{claim:c}
	If $c_2$ is set, then
	$B(c_p, \radius{c_p}) \subseteq B_2(c_2, r)$.
\end{claim}

\begin{proof}
If $|B_2| < \dz$, then $c_p = c_2$ and $r = \radius{c_2}$, 
and hence, $B_2 = B(c_p, \radius{c_p})$. 
When the size of $B_2$ reaches $\dz$, 
the central point $\cp$ moves to the centerpoint of $B_2$,
and $\rc$ is increased by a factor of $(2 + \frac{2}{\alpha})$. 
Because the centerpoint of $B_2$ lies in $B_2$, then $c_p \in B(c_2, \radius{c_2})$.
Thus, if $|B_2| \ge \dz$ then $\len{c_2 c_p} \leq \radius{c_2}$, and therefore,
$$
	\radius{c_p} \leq \frac{2 (\len{c_1 c_2} + \len{c_2 c_p})}{\alpha} \leq \radius{c_2}(1 + \frac{2}{\alpha}).
$$ 
Hence, $B(c_p, \radius{c_p}) \subseteq B_2(c_2, \radius{c_2}(2 + \frac{2}{\alpha}))$.
\end{proof}

\noindent
Now, we prove Invariant (c),
which states $B_1 \cap B_2^* = \emptyset$.

\begin{proof}
A point $p$ can be added to $B_1$ in two cases. % by \Cref{alg:case-1}. 
The first case is in function \textproc{AddToB$_1$}, 
where the point is added to $B_1$ only if it is within distance $\delta$ of the center $c_1$,  
%when $\len{pc_1} \leq \delta$, 
which by invariant (a), 
guaranties $\len{pc_1} < \delta^*$. 
Therefore, $p \not \in B_2^*$ in this case.

The second case is in \Cref{alg:case-1}, when the buffer overflows and $B_2$ is non-empty.
The algorithm takes one of the following actions depending on the size of $B_2$.
If $|B_2| < \dz $, then $c_2 = c_p$.
Algorithm~\ref{alg:case-1} adds $c_2$ to $B_1$.
Suppose by way of contradiction that $c_p \in B_2^*$.
Then, by invariants (b) and (d1), 
$2r^* \leq r \leq \delta/2 < \delta^*/2$. 
Therefore, by Lemma \ref{lem:2r}, there must be at most $z$ points outside $B_1 \cup B_2$,
which contradicts the overflow of the buffer. 
If $|B_2| \geq \dz$, then $c_p$ is the centerpoint of the first $\dz$ points currently in $B_2$. 
In this case, we add all points of $B_2$ to $B_1$. 
By invariant (b), $r \leq \delta/2 < \delta^*/2$. 
Therefore, By Lemma \ref{lem:center-point}, $c_p \in B_1^*$ or $c_p \in B_2^*$. 
Suppose by way of contradiction that $c_p \in B_2^*$. 
In this case, by invariant (d1) and Claim~\ref{claim:c}, % that we proved first,
$B_2$ covers $B(c_p, \radius{c_p})$ and $2r^* \leq r$. 
Therefore, Similar to the previous part, it contradicts the overflow of the buffer. 
\end{proof}
